---
title: 'Observe and check your data in R'
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Check your data in R}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dplyr)
library(tibble)
library(observer)
```


## Create observations from your data with 'observe_if'

The observer package checks that a given dataset passes user-specified rules. The main functions are `observe_if` and `inspect`. 

For instance, according to the documentation of the `diamonds` dataset in package `ggplot2`, 
the column `depth` is equal to `100*2*z/(x+y)`. 
Let us make an observation of this: 

```{r}
df <- ggplot2::diamonds %>% 
  mutate(depth2 = 100*2*z/(x+y)) %>% 
  observe_if(x > 0, 
             y > 0, 
             z > 0, 
             abs(depth-depth2) < 1)

obs(df)
```

We observe that 93 rows fail to satisfy this rule. 
To go further we need to see what is happening; with `inspect` we can select the  rows at stake: 

```{r}
inspect(df, 4)
```

Another way is to write it with standard evaluation: 

```{r}
## Write your predicates first
p <- c(~ x > 0, ~ y > 0, ~ z > 0, 
       ~ abs(depth-depth2) < 1)

## Make observations
df %>% 
  observe_if_(.dots = p) %>% 
  obs()
```



